### 2025-08-22 basic-statistics.qmd
20250822 - clean up code

# Alpha polygon
y.poly <- pmin(y1,y2)
poly1 <- data.frame(x=x, y=y.poly)
poly1 <- poly1[poly1$x >= z_crit, ]
poly1<-rbind(poly1, c(z_crit, 0))  # add lower-left corner

# Beta polygon
poly2 <- df2
poly2 <- poly2[poly2$x <= z_crit,]
poly2<-rbind(poly2, c(z_crit, 0))  # add lower-left corner

# power polygon; 1-beta
poly3 <- df2
poly3 <- poly3[poly3$x >= z_crit,]
poly3 <-rbind(poly3, c(z_crit, 0))  # add lower-left corner

@@

# plot with ggplot2
ggplot(poly, aes(x,y, fill=id, group=id)) +
  geom_polygon(show.legend=F, alpha=I(8/10)) +
  # add line for treatment group
  geom_line(data=df1, aes(x, y, color = "H0", group = NULL, fill = NULL), linewidth = 1.5, show_guide = FALSE) + 
  # add line for treatment group. These lines could be combined into one dataframe.
  geom_line(data=df2, aes(color = "HA", group = NULL, fill = NULL), linewidth = 1.5, show_guide = FALSE) +
  # add vlines for z_crit
  geom_vline(xintercept = z_crit, linewidth=1, linetype="dashed") +
  # change colors
  scale_color_manual(
    "Group",
    values = c("HA" = "#981e0b", "H0" = "black")) +
  scale_fill_manual("test", values= c("alpha" = "#0d6374", "beta" = "#be805e", "power"="#7cecee")) +
  # beta arrow
  annotate("segment", x = 0.1, y = 0.045, xend = 1.3, yend = 0.01, arrow = arrow(length = unit(0.3, "cm")), linewidth = 1) +
  annotate("text", label="beta", x=0, y=0.05, parse=T, size=8) +
  # alpha arrow
  annotate("segment", x = 4, y = 0.043, xend = 3.4, yend = 0.01, arrow = arrow(length = unit(0.3, "cm")), linewidth = 1) +
  annotate("text", label="frac(alpha,2)", x=4.2, y=0.05, parse=T, size=8) +
  # power arrow
  annotate("segment", x = 6, y = 0.2, xend = 4.5, yend = 0.15, arrow = arrow(length = unit(0.3, "cm")), linewidth = 1) +
  annotate("text", label = expression(paste(1-beta, "  (\"power\")")), x = 6.1, y = 0.21, parse = TRUE, size = 8) +
  # H_0 title
  annotate("text", label = "H[0]", x = m1, y = 0.28, parse = TRUE, size = 8) +
  # H_a title
  annotate("text", label = "H[1]", x = m2, y = 0.28, parse = TRUE, size = 8) +
  ggtitle("Statistical Power") +
  # remove some elements
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    panel.background = element_blank(),
    plot.background = element_rect(fill="white"),
    panel.border = element_blank(),
    axis.line = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    plot.title = element_text(size=22))
```

>>>>

# Alpha polygon
y.poly <- pmin(y1,y2)
poly1 <- data.frame(x=x, y=y.poly)
poly1 <- poly1[poly1$x >= z_crit, ]
poly1 <- rbind(poly1, c(z_crit, 0))  # add lower-left corner

# Beta polygon
poly2 <- df2
poly2 <- poly2[poly2$x <= z_crit,]
poly2 <- rbind(poly2, c(z_crit, 0))  # add lower-left corner

# power polygon; 1-beta
poly3 <- df2
poly3 <- poly3[poly3$x >= z_crit,]
poly3 <- rbind(poly3, c(z_crit, 0))  # add lower-left corner

@@

# plot with ggplot2
ggplot(poly, aes(x,y, fill=id, group=id)) +
  geom_polygon(show.legend=F, alpha=I(8/10)) +
  # add line for treatment group
  geom_line(data=df1, aes(x, y, color = "H0", group = NULL, fill = NULL), linewidth = 1.5, show_guide = FALSE) + 
  # add line for treatment group. These lines could be combined into one dataframe.
  geom_line(data=df2, aes(color = "HA", group = NULL, fill = NULL), linewidth = 1.5, show_guide = FALSE) +
  # add vlines for z_crit
  geom_vline(xintercept = z_crit, linewidth=1, linetype="dashed") +
  # change colors
  scale_color_manual(
    "Group",
    values = c("HA" = "#981e0b", "H0" = "black")) +
  scale_fill_manual("test", values= c("alpha" = "#0d6374", "beta" = "#be805e", "power"="#7cecee")) +
  # beta arrow
  annotate("segment", x = 0.1, y = 0.045, xend = 1.3, yend = 0.01, arrow = grid::arrow(length = unit(0.3, "cm")), linewidth = 1) +
  annotate("text", label="beta", x=0, y=0.05, parse=T, size=8) +
  # alpha arrow
  annotate("segment", x = 4, y = 0.043, xend = 3.4, yend = 0.01, arrow = grid::arrow(length = unit(0.3, "cm")), linewidth = 1) +
  annotate("text", label="frac(alpha,2)", x=4.2, y=0.05, parse=T, size=8) +
  # power arrow
  annotate("segment", x = 6, y = 0.2, xend = 4.5, yend = 0.15, arrow = grid::arrow(length = unit(0.3, "cm")), linewidth = 1) +
  annotate("text", label = expression(paste(1-beta, "  (\"power\")")), x = 6.1, y = 0.21, parse = TRUE, size = 8) +
  # H_0 title
  annotate("text", label = "H[0]", x = m1, y = 0.28, parse = TRUE, size = 8) +
  # H_a title
  annotate("text", label = "H[1]", x = m2, y = 0.28, parse = TRUE, size = 8) +
  ggtitle("Statistical Power") +
  # remove some elements
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_blank(),
    panel.background = element_blank(),
    plot.background = element_rect(fill="white"),
    panel.border = element_blank(),
    axis.line = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    plot.title = element_text(size=22))
```

### 2025-08-22 getting-started.qmd
20250822 - clean up code

   install.packages(c(
     "petersenlab","remotes","knitr","rmarkdown","tidyverse","nflverse",
     "tidymodels","easystats","broom","broom.mixed","psych","downlit","xml2",
     "gsisdecoder","progressr","DescTools","pwr","pwrss","WebPower","XICOR",
     "dagitty","ggdag","ggtext","gghighlight","ggExtra","grid","patchwork",
     "pROC", "lme4","lmerTest","MuMIn","emmeans","pbkrtest","sjstats",
     "AICcmodavg","rstan","brms","tidybayes","bbmle","fitdistrplus","sn",
     "mclust","magrittr","viridis","viridisLite","msir","plotly","webshot2",
     "quantmod","fPortfolio","NMOF","nFactors","xts","zoo","forecast","stringi",
     "parallelly","doParallel","missRanger","ggridges","powerjoin","caret",
     "LongituRF","gpboost","mgcv","rms","car","lavaan","lavaanPlot","lavaangui",
     "mice","miceadds","interactions","robustbase","ordinal","MASS",
     "data.table","future","future.apply","SimDesign","domir","GGally","Rglpk",
     "TTR"))
   ```
1. Some necessary packages, including the [`ffanalytics` package](https://github.com/FantasyFootballAnalytics/ffanalytics) [@R-ffanalytics], are hosted in GitHub (and are not hosted on the Comprehensive R Archive Network [CRAN]) and thus need to be installed using the following code (after installing the `remotes` package [@R-remotes] above)[^petersenlabPackageGitHub]:
   ```{r}
   #| eval: false

>>>>

   install.packages(c(
     "petersenlab","remotes","knitr","rmarkdown","tidyverse","nflverse",
     "tidymodels","easystats","broom","broom.mixed","psych","downlit","xml2",
     "gsisdecoder","progressr","DescTools","pwr","pwrss","WebPower","XICOR",
     "dagitty","ggdag","ggtext","gghighlight","ggExtra","patchwork","pROC",
     "lme4","lmerTest","MuMIn","emmeans","pbkrtest","sjstats","AICcmodavg",
     "rstan","brms","tidybayes","bbmle","fitdistrplus","sn","mclust",
     "magrittr","viridis","viridisLite","msir","plotly","webshot2","quantmod",
     "fPortfolio","NMOF","nFactors","xts","zoo","forecast","stringi",
     "parallelly","doParallel","missRanger","ggridges","powerjoin","caret",
     "LongituRF","gpboost","mgcv","rms","car","lavaan","lavaanPlot","lavaangui",
     "mice","miceadds","interactions","robustbase","ordinal","MASS",
     "data.table","future","future.apply","SimDesign","domir","GGally","Rglpk",
     "TTR"))
   ```
1. Some necessary packages, including the [`ffanalytics` package](https://github.com/FantasyFootballAnalytics/ffanalytics) [@R-ffanalytics], are hosted in GitHub (and are not hosted on the Comprehensive R Archive Network [CRAN]) and thus need to be installed using the following code (after installing the `remotes` package [@R-remotes] above)[^petersenlabPackageGitHub]:
   ```{r}
   #| eval: false

### 2025-08-22 index.qmd
20250818 - acknowledgments
20250818 - escape character
20250822 - clean up code

# Automatically create a bib database for R packages
packagesForBibliography <- c(
  "petersenlab", "ffanalytics",
  "knitr", "rmarkdown", "broom", "broom.mixed", "psych", "downlit", "xml2", 
  "gsisdecoder", "progressr", "DescTools", "pwr", "pwrss", "WebPower", "XICOR", 
  "dagitty", "ggdag", "ggtext", "gghighlight", "ggExtra", "grid", "patchwork", 
  "pROC",  "lme4", "lmerTest", "MuMIn", "emmeans", "pbkrtest", "sjstats", 
  "AICcmodavg", "rstan", "brms", "tidybayes", "bbmle", "fitdistrplus", "sn", 
  "mclust", "magrittr", "viridis", "viridisLite", "msir", "plotly", "webshot2", 
  "quantmod", "fPortfolio", "NMOF", "nFactors", "xts", "zoo", "forecast",
  "stringi", "parallelly", "doParallel", "missRanger", "ggridges", "powerjoin", 
  "caret", "LongituRF", "gpboost", "cmdstanr", "mgcv", "rms", "car", "lavaan", 
  "lavaanPlot", "mice", "miceadds", "interactions", "robustbase", "ordinal", 
  "MASS", "data.table", "future", "future.apply", "SimDesign", "domir", 
  "GGally","Rglpk",
  "tidyverse", "ggplot2", "dplyr", "lubridate", # packages that are part of the tidyverse ecosystem
  "nflverse", "nflfastR", "nflseedR", "nfl4th", "nflreadr", "nflplotR", # packages that are part of the nflverse ecosystem
  "tidymodels", "rsample", "parsnip", "recipes", "tune", "yardstick", # packages that are part of the tidymodels ecosysyem
  "easystats", "see", "performance", "parameters", "correlation", "effectsize", # packages that are part of the easystats ecosystem
  "learnr", "remotes", "shiny", "PearsonDS", "lavaangui", "dynamic" # packages that are cited but not used for computation (and thus do not need to be installed by me or readers)
)

@@

This work would not be possible without the important contributions of many people.
This book was made possible by the hard work of the package developers of the `nflverse` ecosystem of R packages, who made football data available, including Sebastian Carl, Ben Baldwin, Lee Sharpe, Tan Ho, and John Edwards.
I thank Dr. Benjamin Motz, who provided consultation and many helpful resources based on his fantasy football statistics class.
I also thank key members of [FantasyFootballAnalytics.net](http://fantasyfootballanalytics.net), including Val Pinskiy, Andrew Tungate, Dennis Andersen, Adam Peterson, Jesse Kartes, and Chris Russo, who helped develop and provide fantasy football-related resources and who helped sharpen my thinking about the topic.
In particular, Andrew Tungate and Dennis Andersen helped develop the [`ffanalytics` `R` package](https://github.com/FantasyFootballAnalytics/ffanalytics) [@R-ffanalytics], which scrapes fantasy football projections from the web and calculates projections averaged across sources.
Andrew Tungate also made important additions to the [`ffanalytics` `R` package](https://github.com/FantasyFootballAnalytics/ffanalytics) [@R-ffanalytics] to allow computing fantasy points given projected (or actual) player statistics, which was necessary for conducting a [bootstrap simulation](#sec-simulationProjectedPointsBootstrapping) of players' projected fantasy points.
I am fortunate to work with such a talented, passionate, and creative team.
I also thank Professor Patrick Carroll, who taught me the value of statistics for answering important questions.
Finally, I thank my wife who, in addition to designing the book cover, is a constant source of love and support.

>>>>

# Automatically create a bib database for R packages
packagesForBibliography <- c(
  "petersenlab", "ffanalytics",
  "knitr", "rmarkdown", "broom", "broom.mixed", "psych", "downlit", "xml2", 
  "gsisdecoder", "progressr", "DescTools", "pwr", "pwrss", "WebPower", "XICOR", 
  "dagitty", "ggdag", "ggtext", "gghighlight", "ggExtra", "patchwork", 
  "pROC",  "lme4", "lmerTest", "MuMIn", "emmeans", "pbkrtest", "sjstats", 
  "AICcmodavg", "rstan", "brms", "tidybayes", "bbmle", "fitdistrplus", "sn", 
  "mclust", "magrittr", "viridis", "viridisLite", "msir", "plotly", "webshot2", 
  "quantmod", "fPortfolio", "NMOF", "nFactors", "xts", "zoo", "forecast",
  "stringi", "parallelly", "doParallel", "missRanger", "ggridges", "powerjoin", 
  "caret", "LongituRF", "gpboost", "cmdstanr", "mgcv", "rms", "car", "lavaan", 
  "lavaanPlot", "mice", "miceadds", "interactions", "robustbase", "ordinal", 
  "MASS", "data.table", "future", "future.apply", "SimDesign", "domir", 
  "GGally","Rglpk","TTR",
  "base","graphics","grDevices","grid","utils", # packages that are included with base R; get the full list with: rownames(installed.packages(priority="base"))
  "tidyverse", "ggplot2", "dplyr", "lubridate", # packages that are part of the tidyverse ecosystem
  "nflverse", "nflfastR", "nflseedR", "nfl4th", "nflreadr", "nflplotR", # packages that are part of the nflverse ecosystem
  "tidymodels", "rsample", "parsnip", "recipes", "tune", "yardstick", # packages that are part of the tidymodels ecosysyem
  "easystats", "see", "performance", "parameters", "correlation", "effectsize", # packages that are part of the easystats ecosystem
  "learnr", "remotes", "shiny", "PearsonDS", "lavaangui", "dynamic" # packages that are cited but not used for computation (and thus do not need to be installed by me or readers)
)

@@

This work would not be possible without the important contributions of many people.
This book was made possible by the hard work of the package developers of the `nflverse` ecosystem of R packages, who made football data available, including Sebastian Carl, Ben Baldwin, Lee Sharpe, Tan Ho, and John Edwards.
I thank Dr. Benjamin Motz, who provided consultation and many helpful resources based on his fantasy football statistics class.
I also thank key members of [FantasyFootballAnalytics.net](http://fantasyfootballanalytics.net), including Val Pinskiy, Andrew Tungate, Dennis Andersen, Adam Peterson, Jesse Kartes, and Chris Russo, who helped develop and provide fantasy football-related resources and who helped sharpen my thinking about the topic.
In particular, Andrew Tungate and Dennis Andersen helped develop the [`ffanalytics` `R` package](https://github.com/FantasyFootballAnalytics/ffanalytics) [@R-ffanalytics], which scrapes fantasy football projections from the web and calculates projections averaged across sources.
Andrew Tungate also made important additions to the [`ffanalytics` `R` package](https://github.com/FantasyFootballAnalytics/ffanalytics) [@R-ffanalytics] to allow computing fantasy points given projected (or actual) player statistics, which was necessary for conducting a [bootstrap simulation](#sec-simulationProjectedPointsBootstrapping) of players' projected fantasy points.
I am fortunate to work with such a talented, passionate, and creative team.
I also thank Professor Patrick Carroll, who taught me the value of statistics for answering important questions.
In addition, a special thanks to those who contributed to the book via comments, suggestions, and pull requests, including George Ellis ([\@MargareeMan](https://github.com/MargareeMan)).
Finally, I thank my wife who, in addition to designing the book cover, is a constant source of love and support.
